---
title: "Assignment 6 - Transportation Problem"
author: "Gordon Wall (gwall2)"
output: github_document
---

##Install and load in relevant packages

```{r}
library(lpSolve)
library(lpSolveAPI)
```

#Formulate Problem

```{r}
trans.model <- make.lp(0,6)
lp.control(trans.model, sense="min")
set.objfn(trans.model, c(622,614,630,641,645,649)) 
add.constraint(trans.model, c(1,1,1,0,0,0), "<=", 100)
add.constraint(trans.model, c(0,0,0,1,1,1), "<=", 120)
add.constraint(trans.model, c(1,0,0,1,0,0), "=", 80)
add.constraint(trans.model, c(0,1,0,0,1,0), "=", 60)
add.constraint(trans.model, c(0,0,1,0,0,1), "=", 70)
```

#Solve Problem

```{r}
solve(trans.model)
```

Success!

#Display OBJFN values

```{r}
get.variables(trans.model)
```

Given the output above, the optimal transportation solution (one that minimizes both shipping and prdouction costs while satisfying the full demand of each warehouse) is 80 units produced from Plant B shipped to Warehouse 1, 60 units produced from Plant A shipped to Warehouse 2, and a combination of 40 units produced from Plant A and 30 units produced from Plant B shipped to Warehouse 3. All three Warehouse monthly demands are met, and both Plants stay within their monthly capacity. This optimal transportation configuration results in a minimized total cost of **$132790**.