---
title: "R Notebook"
author: "Gordon Wall (gwall2)"
output:
  pdf_document: default
  html_notebook: default
---

# loading relevant packages...
```{r, message=FALSE, warning=FALSE, echo=FALSE, results=FALSE}
library(tidyverse)
library(data.table)
library(skimr)
library(arules)
```

# import and examine data...
```{r}
groceries.df <- read.transactions("~/r-directory/grocery-basket/groceries_v2.csv", sep = ",")

summary(groceries.df)
```

# confirm that file was read properly by examining first 10 baskets
```{r}
inspect(groceries.df[1:10])
```

# further examine data; plot of top ten most frequent items in dataset
```{r}
itemFrequencyPlot(groceries.df, topN = 10)
```

##### the transactions in this dataset most frequently sport purchases
##### of whole milk, followed by other vegetables and buns

# train model and extract association rules
```{r}
rules <- apriori(groceries.df, parameter = list(support =0.01, confidence = 0.5))
```


# evaluate model performance
```{r}
summary(rules)
```
##### there have been 15 rules extracted from the data set at 0.01 support and 0.5 confidence
##### the highest lift in this rule set is 3.031

# check top 3 rules by lift
```{r}
inspect(sort(rules, by = "lift")[1:3])
```
##### the number one rule suggests that people who buy citrus fruit and/or root vegetables
##### will also buy an item from the category "other vegetables" as well

















#### References ####
#### https://rpubs.com/emzak208/281776 ####





















