---
title: "Assignment 2 - Carrier Delay Visualization"
author: "Gordon Wall (gwall2)"
output:
  pdf_document: default
  html_notebook: default
---

# workspace set-up
```{r, results=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(caret)
library(skimr)
library(data.table)
library(grid)
library(cowplot)
```

## import relevant data
```{r}
carrier.df <- fread("On_Time_Marketing_Carrier_On_Time_Performance_(Beginning_January_2018)_2018_1.csv")
```

## examine data
```{r}
skim_without_charts(carrier.df)
```
##### A lot of columns have nearly all entries missing
##### these columns provide nothing for analysis and can
##### be removed

##### I will filter for variables that have a complete_rate
##### of less than 1% and create a column names index from this
##### which I will then use to remove them from the data frame

## separate drop-worthy columns
```{r}
drops <- carrier.df %>% skim() %>% dplyr::filter(complete_rate < 0.01) 

drop.index <- drops[,"skim_variable"]

t <- as.vector(drop.index$skim_variable)
```

## drop by drop.index
```{r}
carrier.clean <- select(carrier.df, -t)
```
##### the dataframe is much cleaner now, free of noisey variables

##### for further example, things like year and quarter can be dropped for this
##### analysis since every observation is from january, 2018 (1st quarter)
##### and all entries will be the same in these columns
##### other variables will be dropped with this same logic

## separate drop-worthy columns (second time thru)
```{r}
### dropping columns with more than 600,000 empty data entries 

drops2 <- carrier.clean %>% skim() %>% dplyr::filter(character.empty > 600000)

drop.index2 <- drops2[,"skim_variable"]

q <- as.vector(drop.index2$skim_variable)
```

## drop by second drop.index
```{r}
carrier.clean <- select(carrier.clean, -q)
```

```{r}
### dropping year, month, quarter, day of month columns

carrier.clean <- carrier.clean[,-c("Year", "Quarter", "Month", "DayofMonth")]
```

#### there are variables described in the readme.html
#### file stating that some columns have codes which could've
#### been used for multiple different carriers
#### these destroy the integriy of unique IDs and will be dropped now

```{r}
### dropping columns that start with IATA
### the non-unique ones

carrier.clean <- carrier.clean %>% select(-starts_with("IATA"))
```

```{r}
### dropping other irrelevant/redundant columns

carrier.clean <- carrier.clean[,!c("OriginCityName", "OriginStateName")]

carrier.clean <- carrier.clean[,!c("Duplicate")]

carrier.clean <- carrier.clean[,!c("DestStateName", "DestCityName")]

carrier.clean <- carrier.clean[,!c("Marketing_Airline_Network", "Operated_or_Branded_Code_Share_Partners", "Tail_Number")]
```

## convert to proper variable types
```{r}
carrier.clean$FlightDate <- as.Date(carrier.clean$FlightDate, "%Y-%m-%d")

carrier.clean$DayOfWeek <- as.factor(carrier.clean$DayOfWeek)

carrier.clean$Origin <- as.factor(carrier.clean$Origin)

carrier.clean$OriginState <- as.factor(carrier.clean$OriginState)

carrier.clean$Dest <- as.factor(carrier.clean$Dest)

carrier.clean$DestState <- as.factor(carrier.clean$DestState)

carrier.clean <- carrier.clean %>% rename(operating.airline = `Operating_Airline `)
carrier.clean$operating.airline <- as.factor(carrier.clean$operating.airline)

carrier.clean$DistanceGroup <- as.factor(carrier.clean$DistanceGroup)

carrier.clean$Cancelled <- as.factor(carrier.clean$Cancelled)
carrier.clean$Diverted <- as.factor(carrier.clean$Diverted)

carrier.clean$ArrTimeBlk <- as.factor(carrier.clean$ArrTimeBlk)
carrier.clean$DepTimeBlk <- as.factor(carrier.clean$DepTimeBlk)
```

# Final Check of Clean Dataset
```{r}
skim_without_charts(carrier.clean)
```

# Data Visualization

## Question 1 & 2
##### What is the pattern of arrival traffic and departure traffic delays with respect to days and weeks?
```{r}
subset1a <- carrier.clean %>% 
  select(FlightDate, ArrDelayMinutes) %>% 
  group_by(FlightDate) %>% 
  summarise(mean.arrdelay = mean(ArrDelayMinutes, na.rm = TRUE))

plot1a <- ggplot(subset1a, aes(FlightDate, mean.arrdelay)) +
    geom_bar(stat = "identity", aes(col = mean.arrdelay, fill = mean.arrdelay)) +
    ylab("Avg Arrival Delay (min)") +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          legend.position = "none")
  
subset1b <- carrier.clean %>% 
  select(FlightDate, DepDelayMinutes) %>% 
  group_by(FlightDate) %>% 
  summarise(mean.depdelay = mean(DepDelayMinutes, na.rm = TRUE))

plot1b <- ggplot(subset1b, aes(FlightDate, mean.depdelay, col = mean.depdelay, fill = mean.depdelay)) +
    geom_bar(stat = "identity") +
    ylab("Avg Departure Delay (min)") +
    xlab("Flight Date") +
    theme_minimal() +
    theme(axis.ticks.x = element_line(color = "black"), legend.position = "none")
  
cowplot::plot_grid(plot1a, plot1b, align="v", ncol = 1, rel_heights = c(0.50, 0.50))
```

##### Can you interpret the traffic delays?

######## The graph is of Avg Arrival/Departure Delays across the date range of
######## January 1st thru 31st. The bars outline the avg delay on each day and
######## the x axis ticks section off every 7 days, with each tick being the 
######## start of a new seven-day cycle (week). Interpreting the traffic delays
######## in this format reveal that the month of January, on average, decreases
######## steadily in traffic over the duration of the month, with the middle of the
######## month seeing a few peak high traffic delay spikes in the second and third weeks.
######## Further interpretation could reveal that these spikes have something to do
######## with winter storms in the heart of cold January...

## Question 3
##### Which Airport (‘Origin Airport’) has highest departure delay?
```{r}
subset3 <- carrier.clean %>% 
  select(Origin, DepDelayMinutes) %>% 
  group_by(Origin) %>% 
  summarise(sum.airportdelay = sum(DepDelayMinutes, na.rm = TRUE)) %>%
  arrange(desc(sum.airportdelay)) %>%
  slice(1:25)

options(scipen = 100000)

plot3 <- ggplot(subset3, aes(reorder(Origin, sum.airportdelay), sum.airportdelay)) +
  geom_bar(stat = "identity", aes(col = sum.airportdelay, fill = sum.airportdelay)) +
  coord_flip() +
  ylab("Total January Departure Delay (min)") +
  xlab("Origin Airport") +
  ggtitle("Top 25 Most Delayed Airports on Departure") +
  theme_minimal() +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "none")
  

plot3
```
##### Chicago has the highest departure delay.


## Question 4
##### Which Airport has highest arrival delay?
```{r}
subset4 <- carrier.clean %>% 
  select(Origin, ArrDelayMinutes) %>% 
  group_by(Origin) %>% 
  summarise(sum.airportdelay = sum(ArrDelayMinutes, na.rm = TRUE)) %>%
  arrange(desc(sum.airportdelay)) %>%
  slice(1:25)

plot4 <- ggplot(subset4, aes(reorder(Origin, sum.airportdelay), sum.airportdelay)) +
  geom_bar(stat = "identity", aes(col = sum.airportdelay, fill = sum.airportdelay)) +
  coord_flip() +
  ylab("Total January Arrival Delay (min)") +
  xlab("Origin Airport") +
  ggtitle("Top 25 Most Delayed Airports on Arrival") +
  theme_minimal() +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "none")
  

plot4
```
##### Chicago also has the highest arrival delay.


## Question 5
##### How do you relate the delay pattern to the distance travelled?
```{r}
subset5 <- carrier.clean %>%
  select(DistanceGroup, Distance, ArrDelayMinutes, DepDelayMinutes) %>%
  arrange(DistanceGroup, Distance, ArrDelayMinutes, DepDelayMinutes) %>%
  group_by(DistanceGroup) %>%
  summarise(mean.arrdelay = mean(ArrDelayMinutes, na.rm = TRUE), mean.depdelay = mean(DepDelayMinutes, na.rm = TRUE))

plot5a <- ggplot(subset5, aes(DistanceGroup, mean.arrdelay)) +
  geom_bar(stat= "identity", aes(col = mean.arrdelay, fill = mean.arrdelay)) +
  geom_text(label = round(subset5$mean.arrdelay, 2), col = "white", fontface = "bold", position = position_stack(vjust = 0.5)) +
  ylab("Avg Arrival Delay (min)") +
  ggtitle("Avg Arrival/Departure Delay by Distance Group") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none")

plot5b <- ggplot(subset5, aes(DistanceGroup, mean.depdelay)) +
  geom_bar(stat= "identity", aes(col = mean.depdelay, fill = mean.depdelay)) +
  geom_text(label = round(subset5$mean.depdelay, 2), col = "white", fontface = "bold", position = position_stack(vjust = 0.5)) +
  ylab("Avg Departure Delay (min)") +
  xlab("Distance Group (miles)") +
  theme_minimal() +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "none", axis.text.x = element_text(size = 7, face = "bold")) +
  scale_x_discrete(labels = c("< 250", "250-499", "500-749", "750-999", "1000-1249", "1250-1499", "1500-1749", "1750-1999", "2000-2249", "2250-2499", ">= 2500"))

cowplot::plot_grid(plot5a, plot5b, align="v", ncol = 1, rel_heights = c(0.50, 0.50))
```
##### It appears that distance travelled has a relatively uncorrelated effect
##### on delay, as shown above. I originally speculated that delay would increase
##### with distance, but this was a naive assumption at best. Turns out the highest
##### average delays come from the shortest three-five mileage categories. Maybe
##### this has something to do with the fact that longer flights have the ability to 
##### make up for lost delay time by reaching and maintaing cruising altitude at a faster
##### speed...


## Question 6
##### Is there any correlation between weather delay and carrier delay?
```{r}
subset6 <- carrier.clean %>% 
  select(WeatherDelay, CarrierDelay) %>%
  group_by(WeatherDelay) %>%
  summarise(mean.carrierdelay = mean(CarrierDelay, na.rm = TRUE)) %>%
  na.omit()

plot6 <- ggplot(subset6, aes(WeatherDelay, mean.carrierdelay)) +
  geom_point(stat = "identity", aes(col = mean.carrierdelay, fill = mean.carrierdelay, size = mean.carrierdelay)) +
  ylab("Carrier Delay (min)") +
  xlab("Weather Delay (min)") +
  ggtitle("Carrier Delay vs Delay from Weather") +
  theme_minimal() +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "none")

plot6
```
##### From the above visual, it seems that although some weather delays have impacted
##### carrier delay, the vast majority of data shows that carrier delay has been virtually
##### zero even in the presence of increasing weather delay. You can observe some spikes in
##### carrier delay, which seem random and uncorrelated with weather delay. This dataset would
##### have to be cross-referenced with storm data in the same date-range to determine if recorded
##### heavy storms correlated with the spikes in carrier delay...


## Question 7
##### What is the delay pattern you can find in respective states?
```{r}
subset7 <- carrier.clean %>% 
  select(OriginState, DepDelayMinutes) %>% 
  group_by(OriginState) %>% 
  summarise(sum.statedelay = sum(DepDelayMinutes, na.rm = TRUE)) %>%
  arrange(desc(sum.statedelay))

plot7 <- ggplot(subset7, aes(reorder(OriginState, sum.statedelay), sum.statedelay)) +
  geom_bar(stat = "identity", aes(col = sum.statedelay, fill = sum.statedelay)) +
  coord_flip() +
  ylab("Total January Delay (min)") +
  xlab("Origin State") +
  ggtitle("Total January Delay Classified by Origin Flight State") +
  theme_minimal() +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "none", axis.text.y = element_text(size = 5, face = "bold"))
  

plot7
```
##### It seems from the chart above that delay is typically higher in more
##### trafficked and popular transportation states. The largest delays come from
##### states like California (highest), Illinois (third) and New York (fifth),
##### whereas the lowest delays are sported by more remote states like Arkansas.


## Question 8
##### How many delayed flights were cancelled? (approximation)
```{r}
subset8 <- carrier.clean %>%
  mutate(Delayed = if_else(DepDelayMinutes>0, TRUE, FALSE), Cancelled = if_else(Cancelled == 1, TRUE, FALSE)) %>%
  select(Delayed, Cancelled) %>%
  na.omit()

subset8 <- subset8 %>% 
  group_by(Delayed, Cancelled) %>%
  summarise(total = n()) %>%
  mutate(prop = total/sum(total)) %>%
  filter(Delayed == TRUE)

plot8 <- ggplot(subset8, aes(Cancelled, total)) +
  geom_bar(stat = "identity", position = "dodge", aes(col = Cancelled, fill = Cancelled)) +
  geom_text(label = subset8$total, col = "black", fontface = "bold", position = position_stack(vjust = 0.9)) +
  ylab("Number of Flights") +
  xlab("Cancelled?") +
  ggtitle("Number of Flights Cancelled Due to Delay") +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "right", axis.text.x = element_text(size = 12, face = "bold"))
  
  
plot8
```
##### As shown by the chart above of all delayed flights, only 274 flights
##### were actually cancelled because of the delay. Likely due to the fact that
##### the economy of carrier shipping would rather take a small loss from delay
##### than cancel all together. The people of the USA need their products after all!!


## Question 9
##### How many delayed flights were diverted? (approximation)
```{r}
subset9 <- carrier.clean %>%
  mutate(Delayed = if_else(DepDelayMinutes>0, TRUE, FALSE), Diverted = if_else(Diverted == 1, TRUE, FALSE)) %>%
  select(Delayed, Diverted) %>%
  na.omit()

subset9 <- subset9 %>% 
  group_by(Delayed, Diverted) %>%
  summarise(total = n()) %>%
  mutate(prop = total/sum(total)) %>%
  filter(Delayed == TRUE)

plot9 <- ggplot(subset9, aes(Diverted, total)) +
  geom_bar(stat = "identity", position = "dodge", aes(col = Diverted, fill = Diverted)) +
  geom_text(label = subset9$total, col = "black", fontface = "bold", position = position_stack(vjust = 0.9)) +
  ylab("Number of Flights") +
  xlab("Diverted?") +
  ggtitle("Number of Flights Diverted Due to Delay") +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "right", axis.text.x = element_text(size = 12, face = "bold"))
  
  
plot9
```
##### Similar findings to question 8. Only 640 flights were diverted in the face of delay


## Question 10
##### What time of the day do you find arrival delays?
```{r}
subset10 <- carrier.clean %>%
  select(ArrTimeBlk, ArrDelayMinutes) %>%
  group_by(ArrTimeBlk) %>%
  summarise(mean.arrdelay = round(mean(ArrDelayMinutes, na.rm = TRUE), 4)) %>%
  arrange(ArrTimeBlk)

plot10 <- ggplot(subset10, aes(ArrTimeBlk, mean.arrdelay)) +
  geom_bar(stat="identity", aes(col = mean.arrdelay, fill = mean.arrdelay)) +
  ylab("Avg Arrival Delay") +
  xlab("Time of Day Block") +
  ggtitle("Arrival Delay vs Time of Day") +
  theme_minimal() +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "none", axis.text.x = element_text(size = 8, face = "bold", angle = 45))
  
plot10
```
##### It appears that arrival delay peaks around the end of the 24-hour cycle;
##### within the 8-10 PM range


## Question 11
##### What time of the day do you find departure delays?
```{r}
subset11 <- carrier.clean %>%
  select(DepTimeBlk, DepDelayMinutes) %>%
  group_by(DepTimeBlk) %>%
  summarise(mean.depdelay = round(mean(DepDelayMinutes, na.rm = TRUE), 4)) %>%
  arrange(DepTimeBlk)

plot11 <- ggplot(subset11, aes(DepTimeBlk, mean.depdelay)) +
  geom_bar(stat="identity", aes(col = mean.depdelay, fill = mean.depdelay)) +
  ylab("Avg Departure Delay") +
  xlab("Time of Day Block") +
  ggtitle("Departure Delay vs Time of Day") +
  theme_minimal() +
  theme(axis.ticks.x = element_line(color = "black"), legend.position = "none", axis.text.x = element_text(size = 8, face = "bold", angle = 45))
  
plot11
```
##### Similar findings to question 10.
##### It appears that arrival delay peaks and flattens off
##### within the 7-10 PM range.


































